### 2020-05-01 serial.tex
tweaks to abstract

Generation and serial intervals are under-appreciated --- often treated as auxiliary variables for estimating the reproduction number \RR --- yet important quantities in understanding the outset of an outbreak.
Even though their differences were clearly laid out over a decade ago, there is still a need for clear conceptual and theoretical understanding of their roles in shaping the epidemiological dynamics.
A part of the confusion can be attributed to an apparent paradox.
By definition, both generation and serial intervals describe how infection spreads between individuals, and therefore, one might expect them to potentially provide the same link between the epidemic growth rate $r$ and the reproduction number \RR based on the renewal process.
However, current theory does not explain how two different distributions can give the same link between $r$ and \RR --- some studies have further demonstrated that generation and serial intervals give different $r$--\RR link.
We provide an answer to the paradox by showing that if serial intervals are defined as \emph{forward-looking} intervals during the exponential phase, then the correct link between $r$ and \RR emerges.
We also provide a heuristic way of addressing potential biases that can arise from not accounting for changes in these references and apply our framework to re-analyze serial intervals of COVID-19 collected from China.
Our study further reveals gaps in current practices in outbreak analysis and provides rationale for reassessing serial intervals of COVID-19.

>>>>

Generation and serial intervals are under-appreciated --- often treated as auxiliary variables for estimating the reproduction number \RR --- yet important quantities for charaterizing the outset of an outbreak.
Even though their differences were clearly laid out over a decade ago, there is still a need for clear conceptual and theoretical understanding of their roles in shaping the epidemiological dynamics.
A part of the confusion can be attributed to an apparent paradox.
By definition, both generation and serial intervals describe how infection spreads between individuals, and therefore, one would expect them to potentially provide the same link between the epidemic growth rate $r$ and the reproduction number \RR.
However, current theory does not explain how two different distributions can give the same link between $r$ and \RR --- some studies have further demonstrated that generation and serial intervals give different $r$--\RR link.
We provide an answer to the paradox by showing that if serial intervals are defined as \emph{forward-looking} intervals during the exponential phase, then the correct link between $r$ and \RR emerges.
We also provide a heuristic way of addressing potential biases that can arise from not accounting for changes in these references and apply our framework to serial intervals of COVID-19.
Our study demonstrates the importance of early epidemiological investigation through contact tracing for characterizing initial serial intervals and further provides rationale for reassessing serial intervals of COVID-19.
